---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1** It appears that a total of 17,076 voters became new registrants after Biden's announcement. Montgomery County had the most at 4,847, which makes sense because it is the most populated county in the state.

```{r}
biden_date <- mdy(072124)
total_new_registrants <- md_voters_2024 |>
  filter(
    STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE,
    ymd(STATE_REGISTRATION_DATE) > biden_date
  ) |>
  summarise(
    total_amount = n()
  )
total_new_registrants
```

```{r}
biden_date <- mdy(072124)
county_new_registrants <- md_voters_2024 |>
  group_by(COUNTY) |>
  filter(
    STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE,
    ymd(STATE_REGISTRATION_DATE) > biden_date
  ) |>
  summarise(
    total_amount = n(),
  ) |>
  arrange(desc(total_amount))
head(county_new_registrants)
```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2** There were only three weeks of voter registration after Biden's dropout, and two of those weeks are in the top three for new registration overall. I think we can therefore say that there was definitely a 'boom' of new registrants in the weeks from July 21-August 7.

```{r}
biden_date <- mdy(072124)
md_voters_2024_2 <- md_voters_2024 |>
  mutate(
    COUNTY_REGISTRATION_DATE = ymd(COUNTY_REGISTRATION_DATE),
    week_number = week(COUNTY_REGISTRATION_DATE)
  )
new_voters_per_week <- md_voters_2024_2 |>
  count(week_number) |>
  mutate(
    before_after = case_when(
      week_number > week(biden_date) ~ "post_biden",
      week_number < week(biden_date) ~ "pre_biden"
    )
  ) |>
  arrange(desc(n))
new_voters_per_week
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** The month with the highest number of registrants was July, the same month that Biden dropped out in.

```{r}
new_voters_per_month <- md_voters_2024 |>
  mutate(
    COUNTY_REGISTRATION_DATE = ymd(COUNTY_REGISTRATION_DATE),
    month_number = month(COUNTY_REGISTRATION_DATE)
  ) |>
  count(month_number) |>
  arrange(desc(n))
new_voters_per_month
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** The Capital and Central regions had the most new voters by far, as did Democrats and Unaffiliated voters.

```{r}
new_voters_by_region <- md_voters_2024 |>
  mutate(
    region = case_when(
      COUNTY %in% c("Garrett", "Allegany", "Washington") ~ "Western",
      COUNTY %in% c("Frederick", "Montgomery", "Prince George's") ~ "Capital",
      COUNTY %in% c("Anne Arundel", "Baltimore CIty", "Baltimore County", "Carroll", "Harford", "Howard") ~ "Central",
      COUNTY %in% c("Calvert", "Charles", "Saint Mary's") ~ "Southern",
      COUNTY %in% c("Kent", "Queen Anne's", "Talbot", "Caroline", "Dorchester", "Wicomico", "Somerset", "Worcester", "Cecil") ~ "Eastern Shore"
    )
  ) |>
  count(region) |>
  arrange(desc(n))
new_voters_by_region
```

```{r}
new_voters_by_party <- md_voters_2024 |>
  mutate(
    party_affiliation = case_when(
      PARTY == "REP" ~ "Republican",
      PARTY == "DEM" ~ "Democrat",
      PARTY == "UNA" ~ "Unaffiliated"
    )
  ) |>
  count(party_affiliation) |>
  arrange(desc(n))
new_voters_by_party
```

**Q5** What's an additional question you could ask of this data?

**A5** Why didn't as many Republicans become new registrants after Biden's dropout? That was only a week or so after the Trump assassination attempt, so why didn't the same phenomenon occur for Republicans?
