---
title: "lab_12"
author: "Derek Willis"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**A1.**  The leading word in my dataframe is "client" and I thik the top 10 words describe the main reasons for why Maryland attorneys get sanctioned. Based on the leading words, it seems to most commonly be related to money, which could be misuse of funds or other unethical practices.

```{r}
md_attorney_sanctions <- clean_names(read_csv("data/md_attorney_sanctions.csv"))

unique_words <- md_attorney_sanctions |> select(text) |>
  unnest_tokens(word, text)

data("stop_words")
stop_words <- stop_words |>
  add_row(word="senator") |>
  add_row(word="senators") |>
  add_row(word="maryland") |>
  add_row(word="federal") |>
  add_row(word="u.s") |>
  add_row(word="failing") |>
  add_row(word="md") |>
  add_row(word="project") |>
  add_row(word="failed") |>
  add_row(word="respondent") |>
  add_row(word="engaging") |>
  add_row(word="reprimand") |>
  add_row(word="disbarred") |>
  add_row(word="suspension") |>
  add_row(word="counsel") |>
  add_row(word="justice") |>
  add_row(word="bar") |>
  add_row(word="program")

unique_words |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**A2.** The most common trigram is "attorney trust account" and it appears more than 7% of the time. In legal terms, an attorney trust account is a bank account that holds money on behalf of a client, separate from the law firm's operating funds. As mentioned in the previous question, it seems that misuse of this money may be a common issue in terms of sanctions.

```{r}
md_attorney_sanctions <- md_attorney_sanctions |>
  unnest_tokens(trigram, text, token="ngrams", n=3) |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  filter(!word3 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep=" ")) |>
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)

md_attorney_sanctions
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**A3.** My dataframe output suggests that issues surrounding attorney trust accounts have become more common in recent years, with the highest inclusion of the trigram occurring after 2016. In order to further clarify which reasons are most important, we could perform the same sequence of text analysis but only on the sanctions that include the "attorney trust account" trigram to see the more specific details surrounding the reasoning.

```{r}
top_trigram <- "attorney trust account"

md_sanctions_attorney_trust_account_by_year <- md_attorney_sanctions |>
  filter(str_detect(text, top_trigram)) |>
  group_by(fiscal_year) |>
  summarise(count=n()) |>
  arrange(desc(fiscal_year))

md_sanctions_attorney_trust_account_by_year
```
