---
title: "Class Reference"
author: "Derek Willis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}

setwd("~/jour472/data_journalism_2024_fall")

```

### How to install and load the tidyverse

```{r}
#install.packages("tidyverse")
library(tidyverse)
```

### How to print

```{r}

print("write whatever u want here")

```

### How to assign variables

```{r}

WriteNameOfVariableHere <- 3
#the left-facing arrow assigns whatever u put to a variable with that name
#so we now have a variable called 'WriteNameOfVariableHere' whose value is 3

```

### How to do basic math functions

```{r}

x <- 5
y <- 6
#use + for addition, - for subtraction, * for multiplication, and / for division
#so y+x would return 11

```

### How to find minimum and maximum values

```{r}

#use min() or max()
max(8, 17, 4, 38)
#will return 38

```