---
title: "Class Reference"
author: "Derek Willis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory
The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).
```{r}
setwd("~/Desktop/JOUR472/GitHub/data_journalism_2024_fall")
```

### How to install and load the tidyverse
```{r}
#install.packages("tidyverse")
library(tidyverse)
```

### How to print
```{r}
print("write whatever u want here")
```

### How to assign variables
```{r}
write_name_of_variable_here <- 3
#the left-facing arrow assigns whatever u put to a variable with that name
#so we now have a variable called 'WriteNameOfVariableHere' whose value is 3
```

### How to do basic math functions
```{r}
x <- 5
y <- 6
y+x
17-2
#use + for addition, - for subtraction, * for multiplication, and / for division
#so y+x would return 11
```

### How to find minimum and maximum values
```{r}
#use min() or max()
max(8, 17, 4, 38)
#will return 38
```

### How to read an 'rds' file
```{r}
read_rds(file_name_here)
#will show in Environment
```

### How to get a list of the columns in a dataset
```{r}
glimpse(file_name_here)
#will return a list of the columns, the data type for each column and and the first few values for each column
```

### How to get a preliminary table view of a dataset
```{r}
head(file_name_here)
#will return the first several columns and the first six rows of data in table orientation
```

### How to give a sequence of data commands
```{r}
data |> function
#the '|>' operator tells R to take the data and then do this function to it
```

### Common pattern for analyzing datasets
```{r}
data |> group_by(COLUMN NAME) |> summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))
#summarize() is a very useful dplyr function
#arrange() can arrange your data
```

### Other helpful summarization functions
```{r}
sum() #adds values together
median() #finds the median
mean() #averages values
min() #see line 56 for min and max
max()
```

###How to filter your dataset
```{r}
filter("column_name_here" "!= or ==" 'value name')
#'!=' if you don't want that value included, '==' if you do
```

###How to add a column do your datset
```{r}
mutate("new_column_name_here" = "math to create new column")
#basic math functions apply to the math section
#the name of the new column can be the name of an existing column you want to replace/upgrade
```

###How to add a column with conditions
```{r}
case_when("column_name_here" "< or <= or > or >=" ~ "name_of_value_in_new_column")
#this is a mutate() function
#you can add more conditions to the condition section with "&" for 'and' and "|" for 'or'
#it doesn't just have to be less than, greater than, etc.
```

###How to select specific columns for your codeblock's dataset
```{r}
select("column1_name_here", "column2_name_here", "column3_name_here")
#only the columns you list will show up when you run the codeblock
```

###Lubridate
```{r}
#Lubridate is a tidyverse package that makes dealing with dates, times, etc. much easier

ymd(20101215)
#ymd stands for year, month, day
#will return December 15, 2010, displayed as "2010-12-15"

#dmy, mdy also work
mdy("4/1/17")
#will return April 1, 2017, displayed as "2017-04-01"

bday <- dmy("25/09/2003")
month(bday)
#will return September, displayed as "9"

wday(bday, label = TRUE)
#will return Thursday, displayed as "Thu"

year(bday) <- 2016
wday(bday, label = TRUE)
#will return Sunday, displayed as "Sun", since Sep. 25, 2016 was a Sunday
```

###How to make your column and value names all uppercase
```{r}
str_to_upper("column_or_value_name_here")
#this is a mutate() function
#helpful because R considers 'Baltimore' and 'BALTIMORE' as 2 separate things
#str_to_lower(), str_to_title(), str_to_sentence() also common
```

###How to standardize column names
```{r}
clean_names("name_of_dataset_here")
#makes ever column name lowercase with underscores instead of spaces if applicable
#helpful with inconsistent column names
```

###How to change one column name
```{r}
rename("new_column_name" = "old_column_name")
#don't use quotation marks
```

###How to find duplicate rows in your data
```{r}
get_dupes("name_of_dataset_here")
#will return the original row plus all duplicates
```

###How to change the data type of a column
```{r}
as.numeric("column_name_here")
#as.character() for the other way around
```

###How to include non-duplicate rows in a new dataframe
```{r}
distinct()
#will exclude any duplicate rows
```

###How to titlize all the values in a column
```{r}
str_to_title("column_name_here")
#capitalizes the first letter of each value and lowercases the rest
```

###How to shorten all values in a column
```{r}
str_sub("column_name_here", start="start_number"L, end="end_number"L)
#will reduce every value to only the values between your desired start and end numbers
#for the value "3746283048", a start of 3 and end of 7 would change the value into "46283"
#helpful for reducing 9-digit zip codes to 5
```

###How to detect and correct misspellings in a column
```{r}
str_detect("column_name_here","^misspelling_here") ~ "correct_spelling_here",
    TRUE ~ "column_name_here"
  )
#this is a case_when() function
#second line keeps all the correct spellings as they are
```

###The list() function
```{r}
a = c(1, 2, 3, 4, 5)
b = c("A", "B", "C", "D", "E")
c = c("A", "B", "c", "1", "2")

list(a, b, c)
#exactly what it sounds like, doesn't have to be variables
```

###How to make one datafram from multiple data files
```{r}
bind_rows(list("first_dataframe_here", "second_dataframe_here", "and_so_on"))
#often used with list() because you're pulling from several dataframes
#helpful (and only really applicable) when different dataframes have the same structure (column names, data types, etc.), such as yearly crime reports, election data, or anything that is presented the same way across subsequent datasets
```

###How to join datasets that are not completely alike
```{r}
left_join() #takes all the columns from the left table and only the columns that match in the right one
right_join() #takes all the columns from the right table and only the columns that match in the left one
inner_join() #takes only the columns that are the same across both
full_join() #takes all the columns, generally rare to find a use for
#in the best-case scenario, the two tables we want to join share a common column

join_by() #additional instructions for your join function
#helpful when you have identical info but the column names are different
join_by("column_name_from_first_dataset_here"=="column_name_from_second_dataset_here")
#if join_by() is not included, the join will automatically join by any common columns

#full syntax below:
"first_dataset_here" |> 'type'_join("second_dataset_here", join_by("arguments_here"))
```