---
title: "Class Reference"
author: "Derek Willis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory
The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).
```{r}
setwd("~/Desktop/JOUR472/GitHub/data_journalism_2024_fall")
```

### How to install and load the tidyverse
```{r}
#install.packages("tidyverse")
library(tidyverse)
```

### How to print
```{r}
print("write whatever u want here")
```

### How to assign variables
```{r}
write_name_of_variable_here <- 3
#the left-facing arrow assigns whatever u put to a variable with that name
#so we now have a variable called 'WriteNameOfVariableHere' whose value is 3
```

### How to do basic math functions
```{r}
x <- 5
y <- 6
y+x
17-2
#use + for addition, - for subtraction, * for multiplication, and / for division
#so y+x would return 11
```

### How to find minimum and maximum values
```{r}
#use min() or max()
max(8, 17, 4, 38)
#will return 38
```

### How to read an 'rds' file
```{r}
read_rds(file_name_here)
#will show in Environment
```

### How to get a list of the columns in a dataset
```{r}
glimpse(file_name_here)
#will return a list of the columns, the data type for each column and and the first few values for each column
```

### How to get a preliminary table view of a dataset
```{r}
head(file_name_here)
#will return the first several columns and the first six rows of data in table orientation
```

### How to give a sequence of data commands
```{r}
data |> function
#the '|>' operator tells R to take the data and then do this function to it
```

### Common pattern for analyzing datasets
```{r}
data |> group_by(COLUMN NAME) |> summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))
#summarize() is a very useful dplyr function
#arrange() can arrange your data
```

### Other helpful summarization functions
```{r}
sum() #adds values together
median() #finds the median
mean() #averages values
min() #see line 56 for min and max
max()