---
title: "Reverse Engineering Project"
author: "Cooper, Josh, Jack"
date: "Oct. 6, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates](https://cnsmaryland.org/2022/03/03/millions-in-out-of-state-donations-help-fuel-high-profile-maryland-democratic-governor-candidates/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor) #tidyverse, lubridate, and janitor are the only libraries we need for this project
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

## Analysis 
We had no problems with our data, other than some issues up front. The CNS story never specifically notes what data they're working with, but we gathered (through trial and error) that it was the yearly filing from the beginning of 2022. We had no issues actually working with. In terms of basic data dictionary, we broke each candidate down into a file of their total contribs, titled xx_contribs, and one that was their Maryland contribs, or xx_contribs_md.
```{r}
moore_contribs <- clean_names(read_csv("data/moore_contribs.csv")) #all contributions to Wes Moore

moore_contribs_md <- clean_names(read_csv("data/moore_contribs_md.csv")) #Maryland contributions to Wes Moore

perez_contribs <- clean_names(read_csv("data/perez_contribs.csv")) #all contributions to Tom Perez

perez_contribs_md <- clean_names(read_csv("data/perez_contribs_md.csv")) #Maryland contributions to Tom Perez

king_contribs <- clean_names(read_csv("data/king_contribs.csv")) #all contributions to John King

king_contribs_md <- clean_names(read_csv("data/king_contribs_md.csv")) #Maryland contributions to John King

franchot_contribs <- clean_names(read_csv("data/franchot_contribs.csv")) #all contributions to Peter Franchot

franchot_contribs_md <- clean_names(read_csv("data/franchot_contribs_md.csv")) #Maryland contributions to Peter Franchot
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Among the leading campaigns, some Maryland gubernatorial candidates raised more than half their money from outside the state last year, according to the analysis.”
* **Analysis summary**: Our analysis confirms the results/insinuations of this sentence. We ran code to determine the percentage of the total sum of contributions that came from MD for Moore and Perez (we checked King in question 2), the two candidates mentioned. Our results showed that approximately 42% of Moore's contributions were from MD, 31% for Perez, and 11% for King, which were consistent with the numbers given in the story. In addition, the sums that were given in the story also matched up in our analysis.

```{r}
moore_contribs <- clean_names(read_csv("data/moore_contribs.csv"))
moore_contribs_md <- clean_names(read_csv("data/moore_contribs_md.csv"))
moore_total <- moore_contribs |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all contributions to Moore to find out the total money received
moore_total_md <- moore_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Moore to find out the total money received from inside the state
moore_total
moore_total_md
moore_total_md/moore_total*100 #calculating the percenatge of Moore contribs that came from Maryland
```
```{r}
perez_contribs <- clean_names(read_csv("data/perez_contribs.csv"))
perez_contribs_md <- clean_names(read_csv("data/perez_contribs_md.csv"))
perez_total <- perez_contribs |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all contributions to Perez to find out the total money received
perez_total_md <- perez_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Perez to find out the total money received from inside the state
perez_total
perez_total_md
perez_total_md/perez_total*100 #calculating the percenatge of Perez contribs that came from Maryland
```
### Sentence 2

* **Sentence text**: “John B. King, former secretary of Education under President Barack Obama, raised $2.2 million in campaign funding last year. Only 11%, just under $240,000, came from Maryland donors.”
* **Analysis summary**: Our reverse engineering confirmed this sentence from the story. It showed that King raises approximately $2.2 million, that 10.9% (roughly 11) of it was from MD, and that approximately $240,000 was the amount from Maryland ($238,527 to be exact).

```{r}
king_contribs <- clean_names(read_csv("data/king_contribs.csv"))
king_contribs_md <- clean_names(read_csv("data/king_contribs_md.csv"))
king_total <- king_contribs |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all contributions to King to find out the total money received
king_total_md <- king_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to King to find out the total money received from inside the state
king_total
king_total_md
king_total_md/king_total*100 #calculating the percenatge of King contribs that came from Maryland
```

### Sentence 3

* **Sentence text**: “Moore raised over $1.7 million in Maryland, the highest total of any candidate.”
* **Analysis summary**: The sum of the 2021 year filing for Wes Moore was $1.74 million, which is consistent with the sentence above. We also ran the sums for the other leading candidates (King, Perez, and Franchot), and Moore did indeed have the highest sum from MD. For context, King was roughly $240k, Perez $764k, and Franchot $1.3 million.

```{r}
moore_contribs_md <- clean_names(read_csv("data/moore_contribs_md.csv"))
moore_total_md <- moore_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Moore to find out the total money received from inside the state
moore_total_md
```
```{r} 
king_contribs_md <- clean_names(read_csv("data/king_contribs_md.csv"))
king_total_md <- king_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to King to find out the total money received from inside the state
king_total_md
```
```{r}
perez_contribs_md <- clean_names(read_csv("data/perez_contribs_md.csv"))
perez_total_md <- perez_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Perez to find out the total money received from inside the state
perez_total_md
```
```{r} 
franchot_contribs_md <- clean_names(read.csv("data/franchot_contribs_md.csv"))
franchot_total_md <- franchot_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Franchot to find out the total money received from inside the state
franchot_total_md
```
### Sentence 4

* **Sentence text**: "Moore, King and Perez stand in stark contrast to Peter Franchot, a former Maryland state delegate and current comptroller, who raised over 78% of his $1.66 million from inside Maryland."
* **Analysis summary**: The reverse engineering produced the same findings as the article. Our code found that 78.4% of Franchot's donations were from within the state of Maryland.

```{r}
franchot_contribs <- clean_names(read_csv("data/franchot_contribs.csv"))
franchot_contribs_md <- clean_names(read_csv("data/franchot_contribs_md.csv"))
franchot_total <- franchot_contribs |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all contributions to Franchot to find out the total money received
franchot_total_md <- franchot_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  ) #summing up all Maryland contributions to Franchot to find out the total money received from inside the state
franchot_total
franchot_total_md
franchot_total_md/franchot_total*100 #calculating the percenatge of Franchot contribs that came from Maryland
```

### Sentence 5

* **Sentence text**: "King’s campaign, which raised just under 11% of its money from Maryland, said it received donations from every county in the state..."
* **Analysis summary**: The assertion that King received contributions from every county in the state is false, according to our data. Based on our analysis, King did not receive any contributions from Somerset County.

```{r}
king_md_contribs_by_county <- king_contribs_md |>
  mutate( #creating a new 'county' column based on the address of the contributor. we used str_detect() to create case_when's based on the city name. we used Wikipedia for all of the census-designated places in each county, and then added specific designations for addresses that still hadn't been mutated afterwards. this included a few all-caps spellings, spelling variations (Saint Mary's vs. St. Marys), and very small towns that weren't included on Wikipedia's lists. this may not have been the most efficient way to do this, but it worked!
    county=case_when(
      str_detect(contributor_address, "Barton|Cumberland|Frostburg|Lonaconing|Luke|Midland|Westernport|Bowling Green|Cresaptown|Ellerslie|La Vale|McCoole|Mount Savage|Potomac Park") ~ "Allegany",
      str_detect(contributor_address, "Annapolis|ANNAPOLIS|annapolis|ARNOLD|Harwood|Highland Beach|Annapolis Neck|Arden on the Severn|Arnold|Brooklyn Park|Cape Saint Claire|Crofton|Crownsville|Deale|Edgewater|Ferndale|Hanover|Fort Meade|Friendship|Galesville|Gambrills|Glen Burnie|Herald Harbor|Lake Shore|Linthicum|Maryland City|Davidsonville|Mayo|Naval Academy|Odenton|Parole|Pasadena|Pumphrey|Riva|Riviera Beach|Selby-on-the-Bay|Severn|Severna Park|Shady Side|Millersville|South Gate") ~ "Anne Arundel",
      str_detect(contributor_address,"Baltimore|BALTIMORE|Mount Washington") ~ "Baltimore City",
      str_detect(contributor_address,"Barton|Phoenix|Monkton|Parkton|Glen Arm|Hydes|Sparks|Baldwin|Gwynn Oak|Nottingham|Cumberland|Frostburg|Lonaconing|Luke|Midland|Westernport|Arbutus|Baltimore Highlands|Bowleys Quarters|Carney|Catonsville|Cockeysville|Dundalk|Edgemere|Essex|Garrison|Hampton|Kingsville|Lansdowne|Lochearn|Lutherville|Mays Chapel|Middle River|Milford Mill|Windsor Mill|Overlea|Owings Mills|Parkville|Perry Hall|Pikesville|Randallstown|Reisterstown|Rosedale|Rossville|Timonium|Towson|White Marsh|Woodlawn") ~ "Baltimore County",
      str_detect(contributor_address,"Chesapeake Beach|North Beach|Broomes Island|Calvert Beach|Chesapeake Ranch Estates|Drum Point|Dunkirk|Huntingtown|Long Beach|Lusby|Owings|Prince Frederick|Prnc Frederck|Port Republic|St. Leonard|Saint Leonard|Dowell|Solomons") ~ "Calvert",
      str_detect(contributor_address,"Denton|Federalsburg|Goldsboro|Greensboro|Henderson|Hillsboro|Marydel|Preston|Ridgely|Templeville|Choptank|West Denton|Williston") ~ "Caroline",
      str_detect(contributor_address,"Westminster|Finksburg|Woodbine|Taneytown|Manchester|Mount Airy|New Windsor|Union Bridge|Hampstead|Sykesville|Eldersburg") ~ "Carroll",
      str_detect(contributor_address,"Cecilton|Charlestown|Chesapeake City|Elkton|North East|Perryville|Port Deposit|Rising Sun") ~ "Cecil",
      str_detect(contributor_address,"Indian Head|La Plata|Port Tobacco Village|Bensville|Bryans Road|Bryantown|Cobb Island|Hughesville|Pomfret|Potomac Heights|Rock Point|Saint Charles|Waldorf|Tobacco|White Plains") ~ "Charles",
      str_detect(contributor_address,"Cambridge|Brookview|Church Creek|East New Market|Eldorado|Galestown|Hurlock|Secretary|Vienna|Algonquin|Elliott|Fishing Creek|Madison|Taylors Island") ~ "Dorcester",
      str_detect(contributor_address,"Brunswick|Frederick|Burkittsville|Emmitsburg|Middletown|Mount Airy|Myersville|New Market|Thurmont|Walkersville|Woodsboro|Rosemont|Adamstown|Ballenger Creek|Bartonsville|Braddock Heights|Buckeystown|Green Valley|Jefferson|Libertytown|Linganore|Monrovia|Point of Rocks|Sabillasville|Spring Ridge|Ijamsville|Urbana") ~ "Frederick",
      str_detect(contributor_address,"Accident|Deer Park|Friendsville|Grantsville|Kitzmiller|Loch Lynn Heights|Mountain Lake Park|Oakland|Bloomington|Crellin|Finzel|Gorman|Hutton|Jennings|Swanton") ~ "Garrett",
      str_detect(contributor_address,"Aberdeen|Havre de Grace|Havre De Grace|Abingdon|Edgewood|White Hall|Belcamp|Forest Hill|Bel Air") ~ "Harford",
      str_detect(contributor_address,"Columbia|Glenelg|Lisbon|Woodstock|columbia|Elkridge|Cooksville|Ellicott City|Fulton|Highland|Ilchester|Jessup|North Laurel|Savage|Scaggsville") ~ "Howard",
      str_detect(contributor_address,"Betterton|Chestertown|Galena|Millington|Rock Hall|Butlertown|Edesville|Fairlee|Georgetown|Kennedyville|Still Pond|Tolchester|Worton") ~ "Kent",
      str_detect(contributor_address,"Gaithersburg|Rockville|Takoma Park|Barnesville|Brookeville|Chevy Chase|Chevy Chase View|Chevy Chase Village|Garrett Park|Glen Echo|Kensington|Laytonsville|North Chevy Chase|Poolesville|Somerset|Washington Grove|Martin's Additions|Drummond|Oakmont|Ashton-Sandy Spring|Sandy Spring|Aspen Hill|Bethesda|Brookmont|Burtonsville|Cabin John|Calverton|Clarksburg|Cloverly|Colesville|Damascus|Darnestown|Fairland|Forest Glen|Four Corners|Germantown|Glenmont|Hillandale|Kemp Mill|Layhill|Leisure World|rockville|Montgomery Village|Ashton|North Bethesda|bethesda|North Potomac|Olney|Potomac|Redland|Silver Spring|South Kensington|Travilah|White Oak|Wheaton|GAITHERSBURG|Dickerson|Boyds|TAKOMA PARK|Derwood") ~ "Montgomery",
      str_detect(contributor_address,"Bowie|College Park|District Heights|Glenarden|Greenbelt|Hyattsville|Laurel|Mount Rainier|New Carrollton|Seat Pleasant|Berwyn Heights|Bladensburg|Brentwood|Capitol Heights|Cheverly|Colmar Manor|Cottage City|Cottage city|Eagle Harbor|Edmonston|Fairmount Heights|Forest Heights|Landover Hills|Morningside|North Brentwood|Riverdale Park|University Park|UNIVERSITY PARK|Upper Marlboro|Accokeek|Adelphi|Andrews AFB|Aquasco|Baden|Beltsville|Brandywine|Brock Hall|Brown Station|Calverton|Camp Springs|Cedarville|Chillum|Clinton|Coral Hills|Croom|East Riverdale|Fairwood|Forestville|Fort Washington|Friendly|Glassmanor|Glenn Dale|Hillandale|Hillcrest Heights|Kettering|Konterra|Lake Arbor|Landover|Langley Park|Lanham|Largo|Marlboro Meadows|Marlboro Village|Marlow Heights|Marlton|Maryland Park|Melwood|Mitchellville|National Harbor|Oxon Hill|Peppermill Village|Queensland|Rosaryville|Seabrook|Silver Hill|South Laurel|Springdale|Suitland|HYATTSVILLE|Summerfield|Temple Hills|Walker Mill|West Laurel|Westphalia|Woodlawn|Woodmore|Ft Washington|Riverdale") ~ "Prince George's",
      str_detect(contributor_address,"Barclay|Centreville|Church Hill|Queen Anne|Queenstown|Sudlersville|Chester|Grasonville|Kent Narrows|Kingstown|Stevensville") ~ "Queen Anne's",
      str_detect(contributor_address,"California|Callaway|Charlotte Hall|Golden Beach|Lexington Park|Mechanicsville|Piney Point|St. George Island|Tall Timbers|Wildewood|St Marys City|Saint Inigoes|Leonardtown") ~ "St. Mary's",
      str_detect(contributor_address,"Crisfield|Princess Anne|Chance|Dames Quarter|Deal Island|Eden|Fairmount|Frenchtown-Rumbly|Mount Vernon|Hog's Neck|Cow Island|Jason|Smith Island|West Pocomoke") ~ "Somerset",
      str_detect(contributor_address,"Easton|Oxford|Saint Michaels|Trappe|Doncaster|Dover|York|Wyetown|St Michaels|Cordova|Tilghman Island") ~ "Talbot",
      str_detect(contributor_address,"Hagerstown|Boonsboro|Clear Spring|Funkstown|Hancock|Keedysville|Sharpsburg|Smithsburg|Williamsport|Antietam|Bagtown|Bakersville|Beaver Creek|Big Pool|Big Spring|Breathedsville|Brownsville|Cavetown|Cearfoss|Charlton|Chewsville|Dargan|Downsville|Eakles Mill|Edgemont|Ernstville|Fairplay|Fairview|Fort Ritchie|Fountainhead-Orchard Hills|Gapland|Garretts Mill|Greensburg|Halfway|Highfield-Cascade|Indian Springs|Jugtown|Kemps Mill|Leitersburg|Mapleville|Maugansville|Mercersville|Middleburg|Mount Aetna|Mount Briar|Mount Lena|Paramount-Long Meadow|Pecktonville|Pinesburg|Pondsville|Reid|Ringgold|Robinwood|Rohrersville|Saint James|San Mar|Sandy Hook|Tilghmanton|Trego-Rohrersville Station|Wilson-Conococheague|Yarrowsburg") ~ "Washington",
      str_detect(contributor_address,"Fruitland|Salisbury|Delmar|Hebron|Mardela Springs|Pittsville|Sharptown|Willards|Allen|Bivalve|Jesterville|Nanticoke|Nanticoke Acres|Parsonsburg|Powellville|Quantico|Tyaskin|Waterview|Whitehaven") ~ "Wicomico",
      str_detect(contributor_address,"Fitchburg|Gardner|Leominster|Southbridge|Worcester|Ashburnham|Athol|Auburn|Barre|Berlin|Blackstone|Bolton|Boylston|Brookfield|Charlton|Clinton|Douglas|Dudley|East Brookfield|Grafton|Hardwick|Harvard|Holden|Hopedale|Hubbardston|Lancaster|Leicester|Lunenburg|Mendon|Milford|Millbury|Millville|New Braintree|North Brookfield|Northborough|Northbridge|Oakham|Oxford|Paxton|Petersham|Phillipston|Princeton|Royalston|Rutland|Shrewsbury|Southborough|Spencer|Sterling|Sturbridge|Sutton|Templeton|Upton|Uxbridge|Warren|Webster|West Boylston|West Brookfield|Westborough|Westminster|Winchendon|Athol|Baldwinville|Barre|Brookfield|Clinton|Cordaville|Devens|East Brookfield|East Douglas|Fiskdale|Hopedale|Lunenburg|Milford|North Brookfield|Northborough|Petersham|Rutland|South Ashburnham|South Lancaster|Spencer|Sturbridge|Upton|Warren|Webster|West Brookfield|West Warren|Westborough|Whitinsville|Ocean Pines|Ocean City|Winchendon") ~ "Worcester",
      TRUE ~ contributor_address
    )
  ) |>
  group_by(county) |>
  summarise(
    count=n() #count based on county, hopefully showing at least 1 for each, as that is what the Franchot campaign reported.
  ) |>
  arrange(county)

king_md_contribs_by_county
```

-30-