---
title: "Reverse Engineering Project"
author: "Cooper, Josh, Jack"
date: "Oct. 6, 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates](https://cnsmaryland.org/2022/03/03/millions-in-out-of-state-donations-help-fuel-high-profile-maryland-democratic-governor-candidates/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [“Among the leading campaigns, some Maryland gubernatorial candidates raised more than half their money from outside the state last year, according to the analysis.”]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: “John B. King, former secretary of Education under President Barack Obama, raised $2.2 million in campaign funding last year. Only 11%, just under $240,000, came from Maryland donors.”
* **Analysis summary**: Our reverse engineering did not produce the same findings as the report. According to our data, King's campaign received $2.02 million in 2021, not $2.2 million. Our data also shows $195,000 from Maryland donors, which is 9.7% of the total.

```{r}
king_contribs <- clean_names(read_csv("data/king_contribs.csv"))
king_contribs_md <- clean_names(read_csv("data/king_contribs_md.csv"))
king_total <- king_contribs |>
  summarise(
    total=sum(contribution_amount)
  )
king_total_md <- king_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  )
king_total
king_total_md
king_total_md/king_total*100
```

### Sentence 3

* **Sentence text**: “Moore raised over $1.7 million in Maryland, the highest total of any candidate.”
* **Analysis summary**: The sum of all the 2021 contributions to Wes Moore's campaign is about $1.5 million, which is not exactly what was stated in the article. Thus, this particular finding the reporters specified in the story is unconfirmed. Also, the website does not let you download every single contribution to a gubernatorial campaign in one dataset, so we're also unable to confirm the second part of the statement as of now.

```{r}
moore_contribs_md <- clean_names(read_csv("data/moore_contribs_md.csv"))
moore_total_md <- moore_contribs_md |>
  summarise(
    total=sum(contribution_amount)
  )
moore_total_md
```

### Sentence 4

* **Sentence text**: ["Moore, King and Perez stand in stark contrast to Peter Franchot, a former Maryland state delegate and current comptroller, who raised over 78% of his $1.66 million from inside Maryland."]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: ["King’s campaign, which raised just under 11% of its money from Maryland, said it received donations from every county in the state..."]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-