---
title: "lab_06"
author: "derek willis"
date: "2024-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)
# Load the tidyverse and any other libraries you'll need
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)
```

## Get Our PDF

We'll be working with the [Maryland monthly voter registration update](https://elections.maryland.gov/pdf/vrar/2024_09.pdf) from the State Board of Elections. You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to write a couple of paragraphs that summarize the changes in voter registration in Maryland during the month of September.

## Setup Tabula

Start Tabula, then go to <http://127.0.0.1:8080/> in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds.

This PDF has multiple possible tables to extract. We're going to make four dataframes: 

* new registrations by party
* removals by party
* changes in registration
* current registrations by county.

You will need to draw boxes around four sections and export four CSV files that you will load into R. In Tabula, draw a box around the borders of those tables and click the "Preview & Export Extracted Data" button for each one. The export should look pretty clean, but might be missing some column headers. You'll need to include total columns.

Export each CSV (each one should be called `tabula-2024-09.csv` by default, so you should rename them for clarity to your lab_06/data folder.

From there, you will need to read in and clean up the data, and in particular the headers. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. The pre_lab has an example of how to do this!

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers must not contain spaces and they should have meaningful names, not "x1" or something similar. The current registrations dataframe *MUST* include the county name, but should not include the columns under `Changes` and `Party Affiliation From`. How you do that is up to you (and could involve something outside R), but you can use `select()` with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns.

```{r}
# Party registrations
new_registrants_by_party <- read_csv("data/new_registrants_by_party.csv")
new_registrants_by_party <- clean_names(new_registrants_by_party) |>
  rename(registration_type = x1) |>
  rename(unaffiliated = unaf) |>
  rename(other = oth) |>
  rename(duplicates = dups)
head(new_registrants_by_party)
```

```{r}
# Removals
removals_by_party <- read_csv("data/removals_by_party.csv")
removals_by_party <- clean_names(removals_by_party) |>
  rename(reason = x1) |>
  rename(unaffiliated = unaf) |>
  rename(other = oth)
head(removals_by_party)
```

```{r}
# Changes
registration_changes <- read_csv("data/registration_changes.csv")
registration_changes <- clean_names(registration_changes) |>
  rename(county = x1) |>
  rename(address_change = address) |>
  rename(name_change = name)
head(registration_changes)
```

```{r}
# Current registrations
current_regs_by_county <- read_csv("data/current_regs_by_county.csv")
current_regs_by_county <- clean_names(current_regs_by_county) |>
  rename(county = x1) |>
  rename(unaffiliated = unaf) |>
  rename(other = oth)
current_regs_by_county
```

## Rinse, Repeat: September 2022 data

Do all of that again, for the [September 2022 voter registration update](https://elections.maryland.gov/pdf/vrar/2022_09.pdf), so we can compare the two. If you've done well, this is easy: just do the same steps in Tabula, export the files and use the same code to read in and clean up the data, *just remember to give these dataframes different names* so that you can compare September 2024 to September 2022.

To answer the following questions, you'll need to calculate percentages and do other calculations. Whether you do that by combining or joining dataframes or you write code to run separately on both is up to you.

```{r}
# Party registrations
new_registrants_by_party_22 <- read_csv("data/new_registrants_by_party_22.csv")
new_registrants_by_party_22 <- clean_names(new_registrants_by_party_22) |>
  rename(registration_type = x1) |>
  rename(unaffiliated = unaf) |>
  rename(other = oth) |>
  rename(duplicates = dups)
head(new_registrants_by_party_22)
```

```{r}
# Removals
removals_by_party_22 <- read_csv("data/removals_by_party_22.csv")
removals_by_party_22 <- clean_names(removals_by_party_22) |>
  rename(reason = x1) |>
  rename(unaffiliated = unaf) |>
  rename(other = oth)
head(removals_by_party_22)
```

```{r}
# Changes
registration_changes_22 <- read_csv("data/registration_changes_22.csv")
registration_changes_22 <- clean_names(registration_changes_22) |>
  rename(county = x1) |>
  rename(address_change = address) |>
  rename(name_change = name)
head(registration_changes_22)
```

```{r}
# Current registrations
current_regs_by_county_22 <- read_csv("data/current_regs_by_county_22.csv")
current_regs_by_county_22 <- clean_names(current_regs_by_county_22) |>
  rename(county = x1) |>
  rename(unaffiliated_22 = unaf) |>
  rename(other = oth)
current_regs_by_county_22
```

## Answer questions

Q1. Calculate the percentage of registered Democrats, Republicans and unaffiliated voters for the entire state in September 2024 and September 2022, and then calculate the percentage change for those three party categories (the formula for percentage change is easy to remember: (New-Old)/Old). How would you describe the changes in these partisan categories in a story? What's the most interesting thing about these changes that would be the lede of that story?

A1. The changes for Democrats and Republicans are not that substantial. Democrats saw a 3% drop in registrants while Republicans saw less than a 1% drop. However, unaffliated voters increased by almost 8% from 2022-2024, which is definitely the most interesting thing about all three changes and would be at the top of a story.

```{r}
dem_rep_unaf_regs_24 <- current_regs_by_county |>
  select(county, dem, rep, unaffiliated, total) |>
  mutate(dem_percentage = dem/total * 100) |>
  mutate(rep_percentage = rep/total * 100) |>
  mutate(unaf_percentage = unaffiliated/total * 100)
dem_rep_unaf_regs_22 <- current_regs_by_county_22 |>
  select(county, dem, rep, unaffiliated_22, total) |>
  mutate(dem_percentage = dem/total * 100) |>
  mutate(rep_percentage = rep/total * 100) |>
  mutate(unaf_percentage = unaffiliated_22/total * 100)
reg_changes_22_24 <- dem_rep_unaf_regs_24 |>
  full_join(dem_rep_unaf_regs_22) |>
  filter(county == "TOTAL") |>
  mutate(dem_change = (first(dem_percentage)-second(dem_percentage))/second(dem_percentage)*100) |>
  mutate(rep_change = (first(rep_percentage)-second(rep_percentage))/second(rep_percentage)*100) |>
  mutate(unaf_change = (first(unaf_percentage)-second(unaf_percentage))/second(unaf_percentage)*100)
dem_rep_unaf_regs_24
dem_rep_unaf_regs_22
reg_changes_22_24
```

Q2. Among the Maryland jurisdictions with the most registered voters, which one had the biggest percentage increase in unaffiliated voters between 2022 and 2024? How did that county vote for governor in 2022 and for president in 2020? Contrast the partisan percentages of that county with the county that had the eighth-largest percentage increase in unaffiliated voters. You should display information from the data that shows the counties with the largest percentage increase first.

A2. Prince George's County had the biggest percent increase in unaffiliated voters between 2022 and 2024. PG County voted overwhelmingly in favor of Wes Moore for governor in 2022 and Joe Biden for president in 2020. The county with the 8th-largest increase was Howard County. Both counties are heavily Democratic, though PG County's gap between Dem and Rep is much wider. Both counties have more unaffiliated voters than Republican voters.

```{r}
unaf_regs_by_county_24 <- current_regs_by_county |>
  select(county, unaffiliated, total) |>
  mutate(percentage = unaffiliated/total * 100)
unaf_regs_by_county_22 <- current_regs_by_county_22 |>
  select(county, unaffiliated_22, total) |>
  mutate(percentage_22 = unaffiliated_22/total * 100) |>
  rename(total_22 = total)
unaf_regs_22_24 <- merge(unaf_regs_by_county_22, unaf_regs_by_county_24) |>
  select(county, percentage_22, percentage) |>
  mutate(percent_change = (percentage-percentage_22)/percentage_22*100) |>
  rename(percentage_24 = percentage) |>
  arrange(desc(percent_change))
unaf_regs_22_24
```

Q3. Do the party percentages of new registrants (% Dem, % Rep, % Unaffiliated) from September 2024 and September 2022 match the overall percentage of partisan registration in the state at those times?

In other words, did new registrants mirror the overall partisan composition of the electorate, or were they different? Do any of the methods of registration vary significantly from the overall partisan pattern? Which method is worthy of further exploration?

A3. The percentage of new registrants are very similar to the total registration at those times. Around half were Democrats, which is in-line with overall registration, though unaffiliated voters had a much higher percentage than Republicans, relative to current registration in both years. While no registration type is a true outlier, however Democrats seem to dominate registration by mail, which may be worth further exxploration as to why.

```{r}
new_registrants_by_party <- new_registrants_by_party |>
  select(registration_type, dem, rep, unaffiliated, total) |>
  mutate(dem_percentage = dem/total * 100) |>
  mutate(rep_percentage = rep/total * 100) |>
  mutate(unaf_percentage = unaffiliated/total * 100) |>
  arrange(desc(total))
new_registrants_by_party_22 <- new_registrants_by_party_22 |>
  select(registration_type, dem, rep, unaffiliated_22, total) |>
  mutate(dem_percentage = dem/total * 100) |>
  mutate(rep_percentage = rep/total * 100) |>
  mutate(unaf_percentage = unaffiliated_22/total * 100) |>
  arrange(desc(total))
new_registrants_by_party
new_registrants_by_party_22
```

Q4. Choose your own adventure! Using both the 2024 and 2022 data for one of the dataframes (changes, removals or new registrations), ask and answer a question that makes at least one comparison and requires code to answer. That could involve calculating a percentage, percentage change or the raw difference between two columns.

A4. I did the same process as Q2 but for Republican registrants. Cecil County had, by far, the biggest percent increase in Republican voters between 2022 and 2024. PG County voted slightly in favor of Dan Cox for governor in 2022 and haevily for Donald Trump for president in 2020.

```{r}
rep_regs_by_county_24 <- current_regs_by_county |>
  select(county, rep, total) |>
  mutate(percentage = rep/total * 100)
rep_regs_by_county_22 <- current_regs_by_county_22 |>
  select(county, rep, total) |>
  mutate(percentage_22 = rep/total * 100) |>
  rename(total_22 = total)
rep_regs_22_24 <- merge(rep_regs_by_county_22, rep_regs_by_county_24, by=c("county"), all=TRUE) |>
  select(county, percentage_22, percentage) |>
  mutate(percent_change = (percentage-percentage_22)/percentage_22*100) |>
  #rename(percentage_24 = percentage) |>
  arrange(desc(percent_change))
rep_regs_22_24
```

Q5. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A5. Cecil County is a clear outlier in the percent change in unaffiliated voter registration by county at less than 1% while the next lowest is almost 5%. Why is this? What's going on at Cecil County polls?