---
title: "lab_05"
author: "derek willis"
date: "2024-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
# load the tidyverse
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. This week, we'll revisit our new Maryland voters data and try to join it with contributions to the Harris campaign from Maryland donors made via ActBlue. Can we find new voters who are not registered Democrats who gave to Harris after she became a presidential candidate on July 21? What, if anything, can we say about those people?

We'll be doing something similar to what Ence Morse did for [this Washington Post story](https://www.washingtonpost.com/politics/2024/09/24/kamala-harris-younger-donors/?pwapi_token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJyZWFzb24iOiJnaWZ0IiwibmJmIjoxNzI3MzIzMjAwLCJpc3MiOiJzdWJzY3JpcHRpb25zIiwiZXhwIjoxNzI4NzA1NTk5LCJpYXQiOjE3MjczMjMyMDAsImp0aSI6ImIwYWMxNTk4LWMwMDgtNDE5Ny04MjAyLWVmMWNjZWFmYWZiMiIsInVybCI6Imh0dHBzOi8vd3d3Lndhc2hpbmd0b25wb3N0LmNvbS9wb2xpdGljcy8yMDI0LzA5LzI0L2thbWFsYS1oYXJyaXMteW91bmdlci1kb25vcnMvIn0.3-s0tBkMV7Y9DD21CBsKMlaliGlOKT2egCO-UFT735Y), just on a smaller scale.

First, load two CSV contribution files in the `data` folder, one each for donations made in July and August, and then combine them into a single dataframe.

```{r}
# load the Harris contribution data files and combine them
july_harris_contribs <- read_csv("data/july_harris.csv")
august_harris_contribs <- read_csv("data/august_harris.csv")
harris_contribs <- bind_rows(list(july_harris_contribs, august_harris_contribs))
```

Next, let's read in the new voters (people who registered in 2024):

```{r}
# load the new voters data
new_voters <- read_csv("data/new_voters.csv")
```

Next join the Harris donations to the voter data. Remember, people's names are common, __so you'll want to use name columns and address information__. You'll need to figure out which columns to use from both dataframes. The new_voters data has both a single Address column and various parts of the address, such as HouseNumber. Make a plan for picking your join columns - try out some combinations - and save the results to a new dataframe.

```{r}
# join the combined Harris file to the new voters
new_voters_supporting_harris <- harris_contribs |> 
  inner_join(new_voters, join_by(address_one==Address, first_name==FirstName, last_name==LastName))
# you will need to determine what columns to use

new_voters_supporting_harris
```

## Answer questions

Q1. Explain why you used the columns you did to make the join, and why that result is better than other ways you tried. Are there any problems with the join you made, such as the possibility of false positives?

A1. I used the first, middle, and last name columns from the new_voters file to join with the contributions file, along with the address column. This was the best way to do it because it combined only the necessary info from the voters file.

Q2. Based on your join, how many donations to Harris have come from people who registered this year? How many unique people does that represent, and which donors have given the largest _number_ of donations since July 20? 

A2. There appear to be 365 donations to Harris coming from people who registered this year. These donations are from 270 unique people, with the most contributions (7) being made by Sophia Hoffer-Perkins, Brittany Musolino, and Eric Snipe.

```{r}
new_voter_contribs <- new_voters_supporting_harris |>
  group_by(last_name, first_name) |>
  summarise(
    count=n()
  ) |>
  arrange(desc(count))
new_voter_contribs
```

Q3. What are the party and gender breakdowns of those donors? How many donations has Harris gotten from non-Democrats?

A3. The support for Harris is overwhelmingly Democrat. Only 41 donations came from non-Democrats. Harris received support from men and women, but women had the edge.

```{r}
new_voter_contribs_party <- new_voters_supporting_harris |>
  group_by(Party) |>
  summarise(
    count=n()
  ) |>
  arrange(desc(count))
new_voter_contribs_party
```

```{r}
new_voter_contribs_gender <- new_voters_supporting_harris |>
  group_by(Gender) |>
  summarise(
    count=n()
  ) |>
  arrange(desc(count))
new_voter_contribs_gender
```

Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve just the new voters, or the donors or the joined dataframe, but it cannot be a variation of a previous question (so, for example, you can't ask, "How many Republicans have given to Harris"). First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.

A4: I would be interested to know the geographical trends of the new voters that supported Harris. What counties/towns are home to the majority of these kinds of political participants? I think it is fairly simple: the new_voters_supporting_harris file has a 'city' column so I should just group by that and arrange by the count.

```{r}
new_voters_by_city <- new_voters_supporting_harris |>
  group_by(city) |>
  summarise(
    count=n()
  ) |>
  arrange(desc(count))
new_voters_by_city
```

It seems that outside of Baltimore, cities in Montgomery County are leading the way in this category. Generally, the cities with larger populations are towards the top of this list, which makes sense.